Solutions to Integer Programming from Quantum Annealing

### 2020-08-27 paper/main.tex
first draft

The motivation behind implementing per-qubit level offsets is rooted in the hypothesis that as the problem Hamiltonian is turned on, the system in general will exhibit more disorder. We define disorder by first recasting the QUBO into its Ising form such that the problem Hamiltonian has the form
\begin{equation}
H^{\textrm{Ising}} = \sum_{ij} J_{ij} \sigma^z_i \sigma^z_j + \sum_i h_i \sigma^z_i,
\end{equation}
and recognize that in the limit where $h_i$ is randomly drawn from a Bernoulli distribution we recover the spin-glass model. The spin-glass model enters its glassy state when $|h_i|$ is large, and as a consequence, the wavefunction experiences many-body localization effects, the many-body analogue of Anderson Localization (MBL)~{\color{red} cite some AL papers here}.
\\\\
\noindent\textit{Implementation of annealing offsets}

@@

\subsection{Minor Embedding}
\label{sec:methods:minor_embedding}

@@

@@

>>>>

The motivation behind implementing per-qubit level offsets is rooted in the hypothesis that as the problem Hamiltonian is turned on, the system in general will exhibit more disorder. We define disorder by first recasting the QUBO into its Ising form such that the problem Hamiltonian has the form
\begin{equation}
H^{\textrm{Ising}} = \sum_{ij} J_{ij} \sigma^z_i \sigma^z_j + \sum_i h_i \sigma^z_i,
\end{equation}
and recognize that in the limit where $h_i$ is randomly drawn from a Bernoulli distribution we recover the spin-glass model. The spin-glass model enters its glassy state when $|h_i|$ is large, and as a consequence, the wavefunction experiences many-body localization effects, the many-body analogue of Anderson Localization (MBL)~{\color{red} cite some AL papers here}.
\\\\
\noindent\textit{Implementation of annealing offsets}
\label{sec:discussion:offsets}

@@

\subsection{Minor Embedding}
\label{sec:methods:minor_embedding}
Our proposed offset strategy is motivated by the structure of the Hamiltonian being evaluated by the annealer. As a result, details of the embedding are important. We obtain an embedding for $G(n)$ with the \texttt{embed\_qubo} function provided by the DWave Oceans Python package~\cite{dwave_oceans} under the \texttt{dwave.embedding} module. The same embedding is used for all DWave solves of the same graph (independent of offset), and consequently the simulation solves the resulting embedded Hamiltonian for $G(2)$. Additionally, solving the same graph as a function of offset on the exact same qubits removes (or at least keeps consistent) the systematic effects due to solving a problem on different physical qubits. We note that comparisons between different graphs in Fig.~\ref{fig:baseline} are subject to this uncontrolled systematic.

After embedding the QUBO for $G(2)$ requires 5 qubits (an increase from 4), where by construction, qubits 0 and 3 form the qubit chain. We confirm through brute force evaluation of the eigenvalue decomposition of the 5 qubit Hamiltonian, that the embedding provided by DWave solves the expected ILP problem for $G(2)$, with degenerate ground states at $(0, 1, 0, 0, 0)$ and $(1, 0, 0, 1, 0)$ corresponding to whether vertex 0 or 1 is chosen for the MDS solution, and $(1, 1, 1, 1, 1)$ as the first (non-degenerate) excited state where both vertex 1 and 0 are in the set yielding a valid dominating set but not the minimum dominating set.

The resulting Ising Hamiltonian has external field equal to $h = (2.75, 1.5, -1.0, -1.25, -1.0)$. Following the offset strategy described in Sec.~\ref{sec:discussion:offsets}, qubit(s) 0 (1, 2, 3, 4) are placed in the set with relatively stronger (weaker) final external fields. This imbalance in the two groups may perhaps explain the reason why effects of delaying the weaker fields are more pronounced, since delaying the strong fields only differs from the baseline by a single qubit.

@@

Next, the simulator is applied to the $G(2)$ problem and the result is summarized in Fig.~\ref{fig:td_prob} where the time-dependent overlap with the exact Ising ground state is shown. We observe for all cases that the system undergoes what is analogous to a magnetic phase transition around $s\sim 0.4$. After the phase transition, we are able to confirm that the system collapses to effectively a classical state in the sense that the density matrix becomes a diagonal matrix. The steady increase in probability after the phase transition is a sensitive balance between the competing effects between full-counting statistics and local damping. In our example, full-counting statistics is tuned to be slightly stronger, resulting in what amounts to a final thermal annealing step. If local damping is relatively stronger, then the probability after the phase transition will slowly decrease as the system decoheres into its local ground state. While we believe both effects are important in DWave, the experimental results are not precise enough for us to conclude which is the dominant effect. We emphasize however, that the simulation suggests that the ground state is recovered predominantly due to the quantum phase transition.

@@

Finally, we comment that in order for the simulation to obtain the final state distribution shown in Fig.~\ref{fig:final_state_distribution}, effects of full-counting statistics are required. Absent some dynamical thermalization effect, the $G(2)$ problem is too small such that at a total annealing time of 1$\mu s$, the annealing offsets lift the ground state degeneracy by an amount such that diabatic transitions are absent, and the simulation is populated predominantly by the unique ground state. The final state distribution gives us a very sensitive observable to tune the simulation temperature, and agrees well with the experimental operating temperature.
